Data Migration under a Changing Schema

### 2023-07-14 2022Migration/articleMigrationFACS.tex
Added references to tie 4.3 to 4.2.

   Let $\pair{\dataset}{\schema}$ be the existing system, and let ${\schema'}$ be the schema of the desired system.
   Then we define the migration systems $\migrsys$ and the events as follows:
\begin{enumerate}
\item We take a disjoint union of the relations, defining $\overleftarrow~$ and $\overrightarrow~$ as (overloaded) notation to generate new names, keeping the dataset of the old system and the rules of the new system:
   \begin{align}
   \overleftarrow{\declare{n}{x}{y}}&=\declare{(\id{old}.n)}{x}{y}&\overleftarrow{X} &= \{\overleftarrow{x}\mid x\in X\}\\
   \overrightarrow{\declare{n}{x}{y}}&=\declare{(\id{new}.n)}{x}{y}& \overrightarrow{X} &= \{\overrightarrow{x}\mid x\in X\}\\
   \triples_\migrsys &= \{\triple{a}{\overleftarrow{r}}{b} \mid \triple{a}{r}{b}\in\triples_\dataset\}&
   \instance_\migrsys &= \instance_\dataset\\
   \rels_\theenumi &= \overleftarrow\rels_{\schema} \cup \overrightarrow\rels_{\schema'}
   %&\rules_\theenumi &= \rules_\schema
   \end{align}
\item We create ENFORCE rules to copy equal relations, corresponding to (\ref{extraRelation}) and (\ref{difference}) in Section~\ref{sec:loading}:
   \begin{align}
   \widetilde{\declare{n}{x}{y}} &= \declare{(\id{copied}.n)}{x}{y}\\
   \rels_\theenumi &= \{\widetilde{r}\mid r \in \rels_\schema \cap \rels_{\schema'}\}\\
   \enforces_\theenumi &= \{\overrightarrow{r}\mapsfrom \popF{\overleftarrow{r} - \widetilde{r}} \mid \widetilde{r} \in \rels_2\} \cup \{\widetilde{r}\mapsfrom \popF{\overrightarrow{r}} \mid \widetilde{r} \in \rels_2\}
   \end{align}
\item For each rule, we generate two helper relations, {\tt viols} and {\tt fixed}. Note that Rule~\ref{warnings} only influences the user interface and therefore is not modeled here.
   \begin{align}
   \overleftarrow{\dataset'} &= \pair{\{\triple{x}{{r}}{y}\mid \triple{x}{\overleftarrow{r}}{y} \in \triples_{\dataset'}\}}{\instance_{\dataset'}}\\
   \overrightarrow{\dataset'} &= \pair{\{\triple{a}{{r}}{b} \mid \triple{a}{\overrightarrow{r}}{b}\in\triples_\dataset\}}{\instance_{\dataset'}} \\
   \rels_\theenumi &= \{{\tt viols}_u\mid u \in \rules_{\schema'}\} \cup \{{\tt fixed}_u\mid u \in \rules_{\schema'}\}\\
   \rules_\migrsys &= \{\lambda {\dataset'}.~ \viol{u}{\overrightarrow{\dataset'}} - \pop{{\tt viols}_u - {\tt fixed}_u}{\dataset'} \mid u\in\rules_{\schema'}\}\label{eqn:migrsysrules}\\
   \enforces_{\theenumi} &=\{{\tt viols}_u \mapsfrom \lambda \dataset'.~ \viol{u}{\overleftarrow{\dataset'}}\} \cup
    \{{\tt fixed}_u \mapsfrom \lambda \dataset'.~ {\tt viols}_u - \viol{u}{\overrightarrow{\dataset'}}\}\label{eqn:enforceForRules}
   \end{align}
\item Combining the above into a single migration system:
   \begin{align}
   \concepts_\migrsys &= \concepts_\dataset \cup  \concepts_{\dataset'}\\
   \rels_\migrsys &= \rels_1\cup\rels_2\cup \rels_3\\
   %\rules_\migrsys &= \rules_1\cup \rules_3\\
   \enforces_\migrsys &= \enforces_1\cup\enforces_2\cup\enforces_3
   \end{align}
\end{enumerate}

>>>>

   Let $\pair{\dataset}{\schema}$ be the existing system, and let ${\schema'}$ be the schema of the desired system.
   Then we define the migration systems $\migrsys$ and the events as follows:
\begin{enumerate}
\item We take a disjoint union of the relations, defining $\overleftarrow~$ and $\overrightarrow~$ as (overloaded) notation to generate new names, keeping the dataset of the old system and the rules of the new system:
   \begin{align}
   \overleftarrow{\declare{n}{x}{y}}&=\declare{(\id{old}.n)}{x}{y}&\overleftarrow{X} &= \{\overleftarrow{x}\mid x\in X\}\\
   \overrightarrow{\declare{n}{x}{y}}&=\declare{(\id{new}.n)}{x}{y}& \overrightarrow{X} &= \{\overrightarrow{x}\mid x\in X\}\\
   \triples_\migrsys &= \{\triple{a}{\overleftarrow{r}}{b} \mid \triple{a}{r}{b}\in\triples_\dataset\}&
   \instance_\migrsys &= \instance_\dataset\\
   \rels_\theenumi &= \overleftarrow\rels_{\schema} \cup \overrightarrow\rels_{\schema'}
   %&\rules_\theenumi &= \rules_\schema
   \end{align}
\item We create ENFORCE rules to copy equal relations, corresponding to equation~\ref{extraRelation} through~\ref{fill copies with new.isPartOf} in Section~\ref{sec:loading}:
   \begin{align}
   \widetilde{\declare{n}{x}{y}} &= \declare{(\id{copied}.n)}{x}{y}\\
   \rels_\theenumi &= \{\widetilde{r}\mid r \in \rels_\schema \cap \rels_{\schema'}\}\\
   \enforces_\theenumi &= \{\overrightarrow{r}\mapsfrom \popF{\overleftarrow{r} - \widetilde{r}} \mid \widetilde{r} \in \rels_2\} \cup \{\widetilde{r}\mapsfrom \popF{\overrightarrow{r}} \mid \widetilde{r} \in \rels_2\}
   \end{align}
\item For each rule, we generate two helper relations, {\tt viols} and {\tt fixed}. Note that Rule~\ref{warnings} only influences the user interface and therefore is not modeled here. This corresponds to equation~\ref{violRelation} through~\ref{rule uni}.
   \begin{align}
   \overleftarrow{\dataset'} &= \pair{\{\triple{x}{{r}}{y}\mid \triple{x}{\overleftarrow{r}}{y} \in \triples_{\dataset'}\}}{\instance_{\dataset'}}\\
   \overrightarrow{\dataset'} &= \pair{\{\triple{a}{{r}}{b} \mid \triple{a}{\overrightarrow{r}}{b}\in\triples_\dataset\}}{\instance_{\dataset'}} \\
   \rels_\theenumi &= \{{\tt viols}_u\mid u \in \rules_{\schema'}\} \cup \{{\tt fixed}_u\mid u \in \rules_{\schema'}\}\\
   \rules_\migrsys &= \{\lambda {\dataset'}.~ \viol{u}{\overrightarrow{\dataset'}} - \pop{{\tt viols}_u - {\tt fixed}_u}{\dataset'} \mid u\in\rules_{\schema'}\}\label{eqn:migrsysrules}\\
   \enforces_{\theenumi} &=\{{\tt viols}_u \mapsfrom \lambda \dataset'.~ \viol{u}{\overleftarrow{\dataset'}}\} \cup
    \{{\tt fixed}_u \mapsfrom \lambda \dataset'.~ {\tt viols}_u - \viol{u}{\overrightarrow{\dataset'}}\}\label{eqn:enforceForRules}
   \end{align}
\item Combining the above into a single migration system:
   \begin{align}
   \concepts_\migrsys &= \concepts_\dataset \cup  \concepts_{\dataset'}\\
   \rels_\migrsys &= \rels_1\cup\rels_2\cup \rels_3\\
   %\rules_\migrsys &= \rules_1\cup \rules_3\\
   \enforces_\migrsys &= \enforces_1\cup\enforces_2\cup\enforces_3
   \end{align}
\end{enumerate}
